---
title: "ESM204 Assignment 3"
author: "Simone Albuquerque, Claudia Flores, and Anthony Luna"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyr)
library(tidyverse)
library(janitor)
library(ggplot2)
library(wesanderson)
library(dplyr)
library(scales)
library(purrr)
```

```{r read}
#readin Data
cost_gas_data<- read_csv("assign_3_data.csv") %>% 
  clean_names()
```



```{r lm_setup}
# Linear Regression Low Cost Consumer 
lm_low <- lm(price_dollars ~ q_low_gallons, data=cost_gas_data) 
#print(lm_low)  #  1.169e+01 + -6.611e-05*q
# Linear Regression High Cost Consumer
lm_high <- lm(price_dollars ~ q_high_gallons, data=cost_gas_data) 
#print(lm_high)  #  1.580e+01 + -2.731e-05*q
# Linear Regression Aggregate Cost Consumer
lm_agg <- lm(price_dollars ~ q_aggregrate, data=cost_gas_data) 
#print(lm_agg)  #  1.500e+01 + -2.043e-05*q

```


```{r fun_setup}
# Demand Curve Coefficients based on linear models
I1<-  lm_low$coefficients[1] 
I2<-  lm_high$coefficients[1]
S1<-  lm_low$coefficients[2]
S2<-  lm_high$coefficients[2]

# functional definition of high and low demand based on linear model
low_demand <- function(q)  I1 + (S1*q)
high_demand <- function(q) I2 + (S2*q)

# Determines the 'dominate' curve. This only works for linear models
demand_dom <- function(q){
  if (I1==I2){0}
  else if(I1>I2){low_demand(q)}
  else if(I2>I1){high_demand(q)}
}
# Determines q such that high demand and low demand are equal. Used to find
# the kink in the curve. 
intersect_point <- abs(uniroot(function(q){high_demand(q)-low_demand(q)},
                            interval = c(0,1),extendInt = "yes")$root)

# functional defintion of aggregate demand 
# the function itself. When q is less than the interset, returns the 'dominant'
agg_demand <- function(q) {ifelse(q<intersect_point,demand_dom(q),
                                  lm_agg$coefficients[1] + (lm_agg$coefficients[2]*q))
}

demand_dom_x_intersect <- uniroot(demand_dom,
                                  interval = c(0,1),
                                  extendInt = "yes")$root
# finding the supply curve requires us to find the point at which the agg curve the price is 
# equal to three. We know the current price is untaxed and equal to three.
current_price <- 3
supply_agg_intersect <- uniroot(function(q){agg_demand(q)-current_price}, interval = c(0,1), extendInt = "yes")$root
supply_slope <- current_price/supply_agg_intersect
# functional definition of Supply
supply <- function(q){q*supply_slope} 

#functional definition of environmental costs
local_env_cost_curve<- function(q) 1.5 + (.0000*q) # local MEC $ 1.50
global_env_cost_curve<- function(q) .50 + (.0000*q) # global MEC $.50
total_env_cost_curve<- function(q) 2+ (.0000*q)# total MEC $.50

```

```{r plot_supply_demand}
ggplot()+
  geom_point(data = cost_gas_data, aes(x= q_low_gallons, y = price_dollars),color="red")+
  stat_function(aes(x=0, color = "Low Curve"), fun = low_demand)+ 
  geom_point(data = cost_gas_data, aes(x= q_high_gallons, y = price_dollars),color="blue")+
  stat_function(aes(x=0, color = "High Curve"), fun = high_demand)+
  geom_point(data = cost_gas_data, aes(x= q_aggregrate, y = price_dollars),color="green")+
  stat_function(aes(x=0, color = "Aggregrate Curve"), fun = agg_demand)+
  stat_function(aes(x=0, color = "supply Curve"), fun = supply)  +
  xlim(0,800000)+ 
  scale_y_continuous(breaks=c(0,3,6,9,12,15,18),limits = c(0,18))
```
1. Answer the following Questions:

a. What is the aggregate daily demand curve for gasoline? 

$P=`r round(demand_dom(0),digits=3)``r round(-demand_dom(0)/demand_dom_x_intersect,digits=8)`Q  \text{ when }Q<`r comma(intersect_point)`$

$P=`r round(lm_agg$coefficients[1],digits=3)``r round(lm_agg$coefficients[2],digits=8)`Q\text{ when } Q\geq`r comma(intersect_point)`$


b. What is the supply curve for gasoline? 

$P= `r round(supply_slope,digits=8)`Q$

c. What is the “benefit” to consumers under the status quo?
    
```{r}
# find intersect of aggregate curve at P = $3.00 Q = 587482.43
agg_curve_intersect <- uniroot(function(q){agg_demand(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root

# integrate to find total area under agg demand curve 
gross_consumer_benefit <- integrate(agg_demand,lower=0,upper=agg_curve_intersect)$value

# integrate under the supply curve
producer_benefit <- integrate(supply,lower=0,upper=agg_curve_intersect)$value
  
# subtract gross_consumer and producer to find net consumer benefit
net_consumer_benefit <- gross_consumer_benefit - producer_benefit
```    

Consumer Benefit $=\$`r comma(gross_consumer_benefit)`$

d. What is the “benefit” to producers under the status quo? 

Producer Benefit $=\$`r comma(producer_benefit)`$

e. What is the environmental cost under the status quo (locally and in the rest of the world)?

```{r}
# integrate under each MEC from 0 to the agggrate curve intersect with supply curve
local_env_cost <- integrate(local_env_cost_curve, 0, agg_curve_intersect)$value # 881223.6
global_env_cost <- integrate(global_env_cost_curve, 0, agg_curve_intersect)$value # 293741.2
total_env_cost <-  integrate(total_env_cost_curve, 0, agg_curve_intersect)$value # 1174965
```

Local Environmental Cost  $=\$`r comma(local_env_cost)`$

Global Environmental Cost $=\$`r comma(global_env_cost)`$

Total Environmental Cost  $=\$`r comma(total_env_cost)`$

---

2. How is the current consumer benefit divided between “High” and “Low” income consumers?

```{r}
high_curve_intersect <- uniroot(function(q){high_demand(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root
low_curve_intersect <- uniroot(function(q){low_demand(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root

high_gross_consumer_benefit <- integrate(high_demand,lower=0,upper=high_curve_intersect)$value
low_gross_consumer_benefit <- integrate(low_demand,lower=0,upper=low_curve_intersect)$value
```

High Consumer Benefit $=\$`r comma(high_gross_consumer_benefit)`$

Low Consumer Benefit $=\$`r comma(low_gross_consumer_benefit)`$

---

3. A gas tax of \$1.00/gal. is proposed. What would be the effects of this tax on:
  a. The amount of gasoline produced and consumed.


```{r}
# Defining the tax rate
tax <- 1
#functional supply curve with added tax
supply_tax <- function(q){supply(q)+tax}

#Agg curve intersect with new tax supply curve
agg_tax_curve_intersect <- uniroot(function(q){agg_demand(q)-supply_tax(q)},interval = c(0,1),extendInt = "yes")$root


```

Quantity Produced and consumed $=`r comma(agg_tax_curve_intersect)`$

  b. The price of gasoline.

```{r}
tax_price <- supply_tax(agg_tax_curve_intersect)
```

Quantity Produced and consumed $=`r dollar(tax_price)`$

  c. Welfare of “High” income consumers.
    

```{r}
# New curve intersects for high and low consumers
high_tax_curve_intersect <- uniroot(function(q){high_demand(q)-supply_tax(q)},interval = c(0,1),extendInt = "yes")$root
low_tax_curve_intersect <- uniroot(function(q){low_demand(q)-  supply_tax(q)},interval = c(0,1),extendInt = "yes")$root

# Total consumer benefit under new tax paradigm
high_gross_consumer_welfare_tax <- integrate(high_demand,lower=0,upper=high_tax_curve_intersect)$value
low_gross_consumer_welfare_tax <- integrate(low_demand,lower=0,upper=low_tax_curve_intersect)$value

# Total Consumer taxes, incase the welfare is total benefit minus taxes...
high_tax_total <- high_tax_curve_intersect
low_tax_total <- low_tax_curve_intersect
```

High income consumer welfare $=\$ `r comma(high_gross_consumer_welfare_tax)`$

  d. Welfare of “Low” income consumers.

Low income consumer welfare $\$ =`r comma(low_gross_consumer_welfare_tax)`$

  e. Welfare of gas producers.

```{r}
# Producer welfare
producer_tax_welfare <- integrate(supply,lower=0,upper=agg_tax_curve_intersect)$value
```

Producer consumer welfare $=\$ `r comma(producer_tax_welfare)`$

  f. Local environmental damage.

```{r}
# integrate under each MEC from 0 to the agggrate curve intersect with supply curve
local_env_cost_tax <- integrate(local_env_cost_curve, 0, agg_tax_curve_intersect)$value 
global_env_cost_tax <- integrate(global_env_cost_curve, 0, agg_tax_curve_intersect)$value
total_env_cost_tax<-  integrate(total_env_cost_curve, 0, agg_tax_curve_intersect)$value 
```

Local Environmental Cost with tax  $=\$`r comma(local_env_cost_tax)`$

  g. Rest of world environmental damage.

Global Environmental Cost with tax $=\$`r comma(global_env_cost_tax)`$

  h. Total revenue generated by the tax.
  
```{r}
# multiply the tax by the quantity produced
tax_revenue <- tax*agg_tax_curve_intersect
```

Total tax revenue  $=\$`r comma(tax_revenue)`$

---

4. Now, assume that all revenue from a tax will be redistributed to the two groups in proportion to their pre-tax consumption of gas. For example, if 80% of the gas was consumed by High income consumers, then they get 80% of the tax revenue. Also assume that “Low” income consumers bear all local environmental costs. For a range of gas taxes (ranging from \$0 - \$5.00/gal), calculate the effects of the tax on:
  a. Overall welfare of “High” income consumers

```{r}
# The overall welfare consumers will change such that:
#   - the tax in proportion to pre-tax gas use is added to the welfare
#   - the local environmental cost for the tax based consumption is subtracted from the low income users
# To do this, we need to find:
#   - the proportion of gas consumption of high and low users

low_curve_intersect <- uniroot(function(q){low_demand(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root
high_curve_intersect <- uniroot(function(q){high_demand(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root
low_prop <- low_curve_intersect/agg_curve_intersect 
high_prop <- high_curve_intersect/agg_curve_intersect 

#   - the aggregate quantity consumed for a given tax rate to be used by finding local env cost
#   - the tax revenue for a given tax rate to find the tax redistribution
# 
# To do this lets make a function where tax is the variable t. Then we will construct a dataframe 
# which calculates between the range of 0-5, then we can plot it. We will need a function for both
# the high and low consumers


tax_fun <- function(t) {
  
  #find the new aggregate curve intersect based on the tax
  agg_tax_curve_intersect_var <- uniroot(function(q){agg_demand(q)-(supply(q)+t)},interval = c(0,1),extendInt = "yes")$root
  high_tax_curve_intersect_var <- uniroot(function(q){high_demand(q)-(supply(q)+t)},interval = c(0,1),extendInt = "yes")$root
  low_tax_curve_intersect_var <- uniroot(function(q){low_demand(q)-(supply(q)+t)},interval = c(0,1),extendInt = "yes")$root
  
  #tax times the quantity
  total_tax_revenue <- t*agg_tax_curve_intersect_var
  
  #redistribution of taxes
  redis_low <- low_prop * total_tax_revenue
  redis_high <- high_prop * total_tax_revenue
  
  local_env_cost_tax_var <- integrate(local_env_cost_curve,lower=0,upper=agg_tax_curve_intersect_var)$value
  
  high_welfare_var <- integrate(high_demand,lower=0,upper=high_tax_curve_intersect_var)$value + redis_high
  low_welfare_var <- integrate(low_demand,lower=0,upper=low_tax_curve_intersect_var)$value +redis_low -local_env_cost_tax_var
  producer_welfare_var <- integrate(supply,lower=0,upper=agg_tax_curve_intersect_var)$value
  
  return(data.frame(tax=t,high_welfare = high_welfare_var, low_welfare = low_welfare_var, producer_welfare = producer_welfare_var))
    
  }
  
tax_df <- purrr::map_dfr(seq(from=0.01,to=15,by=0.01),tax_fun)
```


```{r}
ggplot()+
  geom_line(data=tax_df,aes(x=tax,y=producer_welfare,color="Producer Welfare"))+
  geom_line(data=tax_df,aes(x=tax,y=high_welfare,color="High Welfare"))+
  geom_line(data=tax_df,aes(x=tax,y=low_welfare,color="Low Welfare"))

```


  b. Overall welfare of “Low” income consumers
  c. Gas producers

---

Simone

5. A new electric car technology is invented and it lowers the demand curves of all income groups by half (vertically). Under these new demand curves, what are the effects on:
```{r}
# assert new demand curves divided by 50 vertically 
I1_h<-  lm_low$coefficients[1]/2
I2_h<-  lm_high$coefficients[1]/2

low_demand_h <- function(q)  I1_h + S1*q
high_demand_h<- function(q) I2_h + S2*q

```

```{r}

# Determines the 'dominate' curve for new curves. 
demand_dom_h <- function(q){
  if (I1_h==I2_h){0}
  else if(I1_h>I2_h){low_demand_h(q)}
  else if(I2_h>I1_h){high_demand_h(q)}
}
# Determines q such that the new high demand and low demand are equal. #Used to find the kink in the curve. 
intersect_point_h <- abs(uniroot(function(q){high_demand_h(q)-low_demand_h(q)},
                            interval = c(0,1),extendInt = "yes")$root)

# functional defintion of new aggregate demand 
# the function itself. When q is less than the interset, returns the 'dominant'
agg_demand_h <- function(q) {ifelse(q<intersect_point_h,demand_dom_h(q),
                                  lm_agg$coefficients[1]/2 + (lm_agg$coefficients[2]*q))
}

demand_dom_x_intersect_h <- uniroot(demand_dom_h,
                                  interval = c(0,1),
                                  extendInt = "yes")$root
```

Aggregate Curve when Demand is decreased by 50% 

$P=`r round(demand_dom_h(0))``r round(-demand_dom_h(0)/demand_dom_x_intersect_h,digits=6)`Q  \text{ when }Q<`r comma(intersect_point_h)`$

$P=`r round(lm_agg$coefficients[1]/2)``r round(lm_agg$coefficients[2],digits=6)`Q\text{ when } Q\geq`r comma(intersect_point_h)`$



```{r}
# find intersect of high & Low curves at P = $3.00 
high_curve_intersect_h <- uniroot(function(q){high_demand_h(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root

low_curve_intersect_h <- uniroot(function(q){low_demand_h(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root

```
# HELP ANTHONY - IN LINE REFERENCING DRAMA
  a. Gas consumption by “High” income consumers  $=`r comma(high_curve_intersect_h)`$   
  
  b. Gas consumption by “Low” income consumers $=`r comma(low_curve_intersect_h)`$

  c. Gas price
  
```{r}
#New Agg curve intersect with supply curve P = 440637.96
agg_curve_intersect_h <- uniroot(function(q){agg_demand_h(q)-supply(q)},interval = c(0,1),extendInt = "yes")$root
```

```{r}
# integrate under each MEC from 0 to the new agggrate curve intersect with supply curve
local_env_cost_h <- integrate(local_env_cost_curve, 0, agg_curve_intersect_h)$value

global_env_cost_h <-  integrate(global_env_cost_curve, 0, agg_curve_intersect_h)$value 

```

  d. Local environmental damage from gasoline  $=`r comma(local_env_cost_h)`$
  
  e. Rest of world environmental damage from gasoline  $=`r comma(global_env_cost_h)`$

```{r}
# Visualize
ggplot()+
  stat_function(aes(x=0, color = "Low Curve"), fun = low_demand_h)+
  stat_function(aes(x=0, color = "High Curve"), fun = high_demand_h)+
  stat_function(aes(x=0, color = "supply Curve"), fun = supply)+ stat_function(aes(x=0, color = "Aggregrate Curve"), fun = agg_demand_h)+ 
xlim(0,400000)+ 
  scale_y_continuous(breaks=c(0,3,6,9,12,15,18),limits = c(0,10))
```

---

6. Compare two situations to determine what value of T (i.e. what gas tax) makes the local environmental quality equal between these two situations?: 
  a. Gas tax of \$T/gal. but no electric car technology and
  b. No gas tax but with electric car technology. 